---
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# sfnetworks_viz

## Hackathon on `sfnetworks` visualization

This is the issue for this hackathon: https://github.com/luukvdmeer/sfnetworks/issues/7

We started this hackathon with [Pedro Pérez](https://github.com/perezp44) and [John Coene](https://github.com/JohnCoene) by brainstorming on what would be good ways of visualizing `sfnetwork` objects. 

In honor of time, we decided to look into how to use the `sfnetwork` class with [`ggraph`](https://ggraph.data-imaginist.com/index.html).

## `ggraph` mapping

This package is meant to support a grammar of graphics for relational data. It works perfectly with `tbl_graph` objects and since `sfnetworks` subclasses `tbl_graph` we thought it would be not so difficult to already use it inside `ggraph()`. 

### Some background on how `ggraph` works:

- [Layouts](https://ggraph.data-imaginist.com/articles/Layouts.html): Defines the position of the nodes in the plot. 
- [Nodes](https://ggraph.data-imaginist.com/articles/Nodes.html): how the nodes are represented.
- [Edges](https://ggraph.data-imaginist.com/articles/Edges.html): how the edges are represented.

### What we found out:

- We can pass X and Y coordinates to the layout function to place the nodes on the plot. As `sfnetworks` has its nodes as an `sf` object with a `POINT` geometry, we can retrieve those X and Y coordinates easily.
- Nodes will take their placing from the layout, and `geom_node_*()` will basically be in charge of figuring out how it should be represented. Usually we would go for `geom_node_point()` on a "spatial" representation of a network.
- Edges can have different representations between their from and to edges which are already supported when an appropriate layout is established. But, when we want to plot the edges to their faithful location in space, we need to think of a new `geom_edge_*()` function.

### Let's go to the code

We first call the workforce (packages) and save our sample data to an object:

```{r, message = F}
library(ggraph)
library(ggplot2)
library(tidygraph)
library(sfnetworks)
library(sf)
```

```{r}
(roxel_net = as_sfnetwork(st_transform(roxel,31466), directed = F)) 
```

Since `sfnetwork` subclasses `tbl_graph` we can already put it into `ggraph`, but the chosen layout for the nodes might not be the most appropriate:

```{r}
roxel_net %>% 
  ggraph() +
  geom_node_point() +
  geom_edge_link()
```

So, let's create a layout function! As explained before, we can retrieve the X and Y coordinates of the nodes, since they are an `sf` object with `POINT` geometry:

```{r}
layout_sf <- function(graph){
  graph %>% 
    activate("nodes") %>% 
    mutate(
      x = sf::st_coordinates(.)[,"X"],
      y = sf::st_coordinates(.)[,"Y"]
    )
}
```

We can now use this layout inside `ggraph`:

```{r}
roxel_net %>% 
  ggraph(layout = layout_sf) +
  geom_node_point()
```

You will see some spatial distortion up there and that will be due to the CRS present on `sfnetwork` objects. 

So to better plot nodes and edges in their geographical space, we propose creating new `geom_node_*()` and `geom_edge_*()` functions that builds on top of `geom_sf()` in ggplot2 which handles `sf` objects. 

First we fetch the nodes and edges:

```{r}
get_sf_nodes <- function(){
  function(layout) {
    sf::st_as_sf(attr(layout, "graph"), "nodes")
  }
}
get_sf_edges <- function(){
  function(layout) {
    sf::st_as_sf(attr(layout, "graph"), "edges")
  }
}
```

And then we can create `geom_node_sf()` and `geom_edge_sf()`:

```{r}
geom_node_sf <- function (mapping = NULL, data = get_sf_nodes(), ...){
  geom_sf(data = data, mapping = mapping, ...)
}
geom_edge_sf <- function (mapping = NULL, data = get_sf_edges(), ...){
  geom_sf(data = data, mapping = mapping, ...)
}
```

And we can easily add it to our plot!

```{r}
roxel_net %>% 
  ggraph(layout = layout_sf) +
  geom_node_sf(color = 'grey30', size = 0.5) +
  geom_edge_sf(aes(color = type)) 
```

With any edge representation we want:

```{r}
roxel_net %>% 
  ggraph(layout = layout_sf) +
  geom_node_sf(color = 'grey30', size = 0.5) +
  geom_edge_arc(aes(color = type)) 
```

And there we have it! Note how the distortion get's handle by `geom_sf()` very nicely!

### (Update of) open issues:

- We still need to think how to handle lat/long data on the layout_sf function, since unprojected networks can appear distorted when plotted. 
  + Creating a geom_node_sf as well as a geom_edge_sf would solve the problem, since it is not only for lat/long or unprojected networks.
- We should take a look internally at `layout_tbl_graph_manual` to implement the new layout.
  + It could be that the CRS should be handled on the layout function. Otherwise, when plotting nodes or edges without teh new `sf` geoms, the network is distorted.  

### Other visualization opportunities:

- explore 3D representation, maybe with `rayshader` 
  + this can take a look into st_zm coordinates
- maybe look into interactive plotting: `mapview`

## Summary:

- A `layout_sf()` function can be implemented to plot the graph in their geographical space. 
- Two new functions: `geom_node_sf()` and `geom_edge_sf()` functions can be implemented to plot the graph in its geographical space and still access their columns to map aesthetics.

The ultimate goal would be to create a pull request to `ggraph`!
